---
title: "5 ReprEx tips for getting code help"
subtitle: "HARUG! #79"
author: "Ed Harris"
date: "2022-01-26"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What is ReprEx?

ReprEx means **Reproducible Example**

When asking for help, often online, it helps to provide enough information to EXACTLY REPRODUCE any error you might be getting

Also, take care to provide complete information

- **description of what you are trying to do**

- **minimal data example**

- **code you have already tried** using minimal data

## Is this ReprEx?

Q: Hello I'm adding legends to my boxplots but I found that if I clear the global environment this line

 legend(x = 0.5, y = 28, legend = c('Year 17', 'Year 18'),
              fill = c(3, 5)))

stops working, it adds the legend correctly if I run the other lines and generate the graph beforehand but if I run the lot together I get the error message

Error in strwidth(legend, units = "user", cex = cex, font = text.font) :
  plot.new has not been called yet

It would seem untidy to me to leave the script for the legend separate. Is there a way to get it to run with the rest of the code for that graph please?

## Answer to a non-Reprex:

A: I would need a bit more code about what you are trying to answer, and your output. 

I do not understand what scenario would make you clear your global environment between making a plot and drawing a legend on your plot?

e.g. something like this reproducible example works fine:

## Answer to a non-Reprex:

```r
var <- c(1,2,3,4)
factor <- c('a','a','b','b')
data <- data.frame(var, factor)

boxplot(var~factor, data=data,
        col=c('red','blue'))

legend(x=1, y = 3.5, legend = c('a', 'b'),
       col=c('red','blue'), pch=15)
```    

## Answer to a non-Reprex:

```{r}
var <- c(1,2,3,4)
factor <- c('a','a','b','b')
data <- data.frame(var, factor)
boxplot(var~factor, data=data,
        col=c('red','blue'))
legend(x=1, y = 3.5, legend = c('a', 'b'),
       col=c('red','blue'), pch=15)

```


## A good ReprEx has 5 features:

- context

- what do you want to do?

- what have you tried (code, not screenshot)?

- minimal data (probably also code)

- error message or wrong outcome


## Example: Compare means of just 2 of 3 flower spp

```{r, echo=T}
data(iris)
head(iris)

```


## Example: Compare means of just 2 of 3 flower spp

```r
my_iris <- iris[which(iris$Species != 'virginica')]

Error in `[.data.frame`(iris, which(iris$Species != "virginica")) : 
  undefined columns selected

```

## Go through the steps

- context

- what do you want to do?

- what have you tried (code, not screenshot)?

- minimal data (probably also code)

- error message or wrong outcome

## references

[Hadley Wickham on ReprEx](https://gist.github.com/hadley/270442#file-reproducible-md)

[Stack Overflow speaks](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example)
