---
title: "Alistair's beet story"
author: "Ed Harris"
date: "2021-11-17"
output: pdf_document
---

```{r setup, include=FALSE, warning = F}
knitr::opts_chunk$set(echo = FALSE)
```

## 00 Background

**The replication is too small to perform 2-way ANOVA** legitimately in my opinion (best practice is 3 reps per main effect). E.g., you would be advised to have a minimum of n=6 reps for 2-way ANOVA and no interaction effect, but here you have n=3.  

**The design does not really fit a dose design** based on the information I have (not enough dose levels or replication).

**I would suggest 1-way ANOVA** converting the 2 factors to a single one with 5 levels: control, A100, A200, B100, B200. With this analysis, post hoc tests would allow further mean comparisons, but the statistical power will probably be very low.

```{r, warning=F}
library(openxlsx)
library(agricolae)
data <- read.xlsx('Alistair-data-ed.xlsx')
data$treat2 <- factor(data$treat2, 
                      levels = c('Control', 'A100', 
                                 'A200', 'B100', 'B200'))
```

\newpage

## **Plot_wt_g** 1-way ANOVA and *post hoc*

```{r, warning = F}
boxplot(Plot_wt_g ~ treat2, data = data,
        xlab = 'Treatment',
        ylab = 'Sample weight (g)')
stripchart(Plot_wt_g ~ treat2, data = data,
           add=T, vert = T, method = 'jitter',
           pch = 16, col = 'red')


m1 <- aov(Plot_wt_g ~ treat2, data = data)
summary(m1)
TukeyHSD(m1)
# HSD.test(y = m1, trt = 'treat2', group=T, console=T)

```

\newpage

## **Sugar_pct** 1-way ANOVA and *post hoc*

```{r, warning = F}

boxplot(Sugar_pct ~ treat2, data = data,
        xlab = 'Treatment',
        ylab = 'Sugar %')
stripchart(Sugar_pct ~ treat2, data = data,
           add=T, vert = T, method = 'jitter',
           pch = 16, col = 'red')


m2 <- aov(Sugar_pct ~ treat2, data = data)
summary(m2)
TukeyHSD(m2)

```

\newpage

## **AmNit** 1-way ANOVA and *post hoc*

```{r, warning = F}

boxplot(AmNit ~ treat2, data = data,
        xlab = 'Treatment',
        ylab = expression(paste('NH'[4],'NO'[3])))
stripchart(AmNit ~ treat2, data = data,
           add=T, vert = T, method = 'jitter',
           pch = 16, col = 'red')


m3 <- aov(AmNit ~ treat2, data = data)
summary(m3)
TukeyHSD(m3)

```

\newpage

## **Sod** 1-way ANOVA and *post hoc*

```{r, warning = F}

boxplot(Sod ~ treat2, data = data,
        xlab = 'Treatment',
        ylab = 'Na (mg/100g)')
stripchart(Sod ~ treat2, data = data,
           add=T, vert = T, method = 'jitter',
           pch = 16, col = 'red')


m4 <- aov(Sod ~ treat2, data = data)
summary(m4)
TukeyHSD(m4)

```

\newpage

## **Pot** 1-way ANOVA and *post hoc*

```{r, warning = F}

boxplot(Pot ~ treat2, data = data,
        xlab = 'Treatment',
        ylab = 'K (mg/100g)')
stripchart(Pot ~ treat2, data = data,
           add=T, vert = T, method = 'jitter',
           pch = 16, col = 'red')


m5 <- aov(Pot ~ treat2, data = data)
summary(m5)
TukeyHSD(m5)

```
